**The status of the microbial census: an update**

### 2015-11-20 Schloss_Census2_mBio_2015.Rmd
Add text/code for taxonomic analysis

arch <- read.table(file='data/process/archaea.v123.metadata', header=T, row.names=1, stringsAsFactors=FALSE)
arch_year_deposited <- gsub("^(\\d\\d\\d\\d)-.*", "\\1", arch$date)
```
\vspace{35mm}

@@

	b_is_cult[rownames(bact) == 'AJ224540.MxeCultu'] <- FALSE
#	aquifer1 - scrap
#	bact[bact_phylum=="Candidate_division_OP3" & !is.na(bact$isolate),] - AJ224540.MxeCultu

@@

cult_otus <- unique(bact$otu[b_is_cult])			# otus with a cultured representative
b_in_cult_otu <- bact$otu %in% cult_otus

#total number of OTUs by phylum
n_all_otus_phylum <- aggregate(bact$otu, by=list(bact_phylum), function(x){length(unique(x))})

#total number of OTUs with a cultured rep by phylum
n_cult_otus_phylum <- aggregate(bact$otu[b_in_cult_otu], by=list(bact_phylum[b_in_cult_otu]), length)

#need to merge these two tables...

@@

* % of OTUs with a cultured representative by phylum
	- basically done.

* do these new cultures represent new OTUs?

***The census is taxonomically skewed***
* For which phyla has there been substantial increases in sampled diversity since 2004?

***New technologies to access novel biodiversity***
  - Single cell genomics

>>>>

bact$category[bact$category=="ZI" & !is.na(bact$category)] <- "ZA"

arch <- read.table(file='data/process/archaea.v123.metadata', header=T, row.names=1, stringsAsFactors=FALSE)
arch_year_deposited <- gsub("^(\\d\\d\\d\\d)-.*", "\\1", arch$date)

arch$category[arch$category=="BDBD" & !is.na(arch$category)] <- "BD"

```
\vspace{35mm}

@@

	b_is_cult[rownames(bact) == 'AJ224540.MxeCultu'] <- FALSE
#	aquifer1 - scrap
#	bact[bact_phylum=="Candidate_division_OP3" & !is.na(bact$isolate) & !is.na(bact_phylum),] - AJ224540.MxeCultu

@@

	b_cult_otus <- unique(bact$otu[b_is_cult])
	b_in_cult_otu <- bact$otu %in% b_cult_otus

	n_all_otus_phylum <- aggregate(bact$otu, by=list(bact_phylum), function(x){length(unique(x))})
	n_cult_otus_phylum <- aggregate(bact$otu[b_in_cult_otu], by=list(bact_phylum[b_in_cult_otu]), function(x){length(unique(x))})

	b_otus_phylum <- data.frame(matrix(rep(0, nrow(n_all_otus_phylum)*2), ncol=2, dimnames=list(n_all_otus_phylum$Group.1, c('cultured', 'total'))))

	b_otus_phylum[n_all_otus_phylum$Group.1,"total"] <- n_all_otus_phylum$x
	b_otus_phylum[n_cult_otus_phylum$Group.1,"cultured"] <- n_cult_otus_phylum$x

	a_cult_otus <- unique(arch$otu[a_is_cult])
	a_in_cult_otu <- arch$otu %in% a_cult_otus

	n_all_otus_phylum <- aggregate(arch$otu, by=list(arch_phylum), function(x){length(unique(x))})
	n_cult_otus_phylum <- aggregate(arch$otu[a_in_cult_otu], by=list(arch_phylum[a_in_cult_otu]), function(x){length(unique(x))})

	a_otus_phylum <- data.frame(matrix(rep(0, nrow(n_all_otus_phylum)*2), ncol=2, dimnames=list(n_all_otus_phylum$Group.1, c('cultured', 'total'))))

	a_otus_phylum[n_all_otus_phylum$Group.1,"total"] <- n_all_otus_phylum$x
	a_otus_phylum[n_cult_otus_phylum$Group.1,"cultured"] <- n_cult_otus_phylum$x

	b_n1_cult_otus <- sum(table(bact$otu[b_in_cult_otu])==1)/length(table(bact$otu[b_in_cult_otu]))
	a_n1_cult_otus <- sum(table(arch$otu[a_in_cult_otu])==1)/length(table(arch$otu[a_in_cult_otu]))
```

@@

* % of OTUs with a cultured representative by phylum and in aggregate
	- b_otus_phylum
	- a_otus_phylum

* do these new cultures represent new OTUs by phylum and in aggregate
	- b_n1_cult_otus
	- a_n1_cult_otus

* need a plot showing % cultured sequences and % cultured OTUs

***The census is taxonomically skewed***

```{r taxon_census}
	b_phylum_year <- table(bact_year_deposited, bact_phylum)
	b_n_phylum_2015 <- ncol(b_phylum_year)
	b_phylum_2006 <- b_phylum_year[rownames(b_phylum_year) %in% as.character(1983:2006),]
	b_n_phylum_2006 <- sum(apply(b_phylum_2006, 2, sum) > 0) #how many phyla were there?
	b_n_old_seqs <- sum(b_phylum_2006)						#how many seqs in 2006?
	b_n_old_seqs_phyla <- apply(b_phylum_2006, 2, sum)		#how many sequences in each phyla?

	b_new_phyla <- colnames(b_phylum_year)[apply(b_phylum_2006, 2, sum) == 0] #what are the new phyla?
	b_new_phyla_names <- paste(paste(b_new_phyla[1:(length(b_new_phyla)-1)], collapse=", "),
							 b_new_phyla[length(b_new_phyla)], sep=", and ")

	a_phylum_year <- table(arch_year_deposited, arch_phylum)
	a_n_phylum_2015 <- ncol(a_phylum_year)
	a_phylum_2006 <- a_phylum_year[rownames(a_phylum_year) %in% as.character(1983:2006),]
	a_n_phylum_2006 <- sum(apply(a_phylum_2006, 2, sum) > 0) #how many phyla were there?
	a_n_old_seqs <- sum(a_phylum_2006)						#how many seqs in 2006?
	a_n_old_seqs_phyla <- apply(a_phylum_2006, 2, sum)		#how many sequences in each phyla?

	a_new_phyla <- colnames(a_phylum_year)[apply(a_phylum_2006, 2, sum) == 0] #what are the new phyla?
	a_new_phyla_names <- paste(a_new_phyla, collapse=" and ")
	a_new_phyla_names <- gsub("_", " ", a_new_phyla_names)
	a_new_phyla_names <- gsub("\\(\\S*\\)", "", a_new_phyla_names)
```

* For which phyla has there been substantial increases in sampled diversity since 2004?
	- new phyla: b_new_phyla_names, a_new_phyla_names
	- Figure phylum_effort.pdf

* generate a heatmap of category by phylum relative abundance
	- Figure phylum_category_heatmap

***New technologies to access novel biodiversity***
  - Single cell genomics
  - Metagenomes
